### 2011-07-27 doc/theses/svirec/kap2.tex
added some new examples

@@

Before defining of the XML integrity constraint, let us introduce some notation used in the definition. A \emph{path atom} is an expression of the form $[x_1]p[x_2]$, where $p$ is a path expression, $x_1$ and $x_2$ are terms, and $x_1 \not \in Str$.\\
A conjunction of path and built-in atoms $C = [X_1]p_1[Y_1] \cap \dots \cap [X_n]p_n[Y_n] \cap U_1\theta_1 V_1 \cap \dots \cap U_k \theta_k V_k$ is said to be \emph{safe} if all variables in $C$ are \emph{range restricted}, i.e. if
\begin{itemize}
 	\item for every $[X_i]p_i[Y_i]$, either $X_i$ is a constant (node identifier of a string), or there is some $[X_j]p_j[Y_j]$ in $C$ where $X_j$ is range restricted;
    \item for every built-in term $U_i\theta_i V_i$ occurring in $C$, if $\theta_i$ is equal to $"="$ then at least one of the two terms is range restricted; otherwise both $U_i$ and $V_i$ must be range restricted.
 \end{itemize}
A rootless tree formula is an expression of the form $p(\Phi_1 \land \dots \land \Phi_k)$ where $\Phi_i$ is a rootless path formula (expression of the form $p[y]$ where $p$ is a path expression and $y$ is a term) or a rootless tree formula and $p$ is a path expression. A tree atom is an expression of the form $[x]T$ where $T$ is a rootless tree formula and $x$ is a term.

@@

@@

\begin{define}[Functional Dependency]\label{fd1}
Being given a DTD $D$, a {\sl functional dependency} on $D$ is an expression of the form $S \rightarrow p$, where S is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
\end{define}

@@

>>>>

\begin{example}\label{pathExample}
Consider the XML tree $XT$ from Example \ref{example1ref} conforming the DTD $D$ defined below, which is representing a collection of books.
\begin{verbatim}
<!ELEMENT bib (book+)>
<!ELEMENT book (written_by, title)>
<!ELEMENT written_by (author+)>
<!ELEMENT author (name)>
<!ATTLIST author ano CDATA>
<!ELEMENT name PCDATA>
<!ELEMENT title PCDATA>
\end{verbatim}

The set $paths(D)$ contains the following paths:
\begin{align}
paths(D) &= \left\{/bib, /bib/book, /bib/book/written\_by,\right.\nonumber\\
&\qquad \left. /bib/book/written\_by/author, \right.\nonumber\\
&\qquad \left. /bib/book/written\_by/author/name, \right.\nonumber\\
&\qquad \left. /bib/book/written\_by/author/name/S, \right.\nonumber\\
&\qquad \left. /bib/book/written\_by/author/@ano, \right.\nonumber\\
&\qquad \left. /bib/book/title\nonumber\right\}
\end{align}\qed
\end{example}

@@

Before defining of the XML integrity constraint, let us introduce some notation used in the definition. A \emph{path atom} is an expression of the form $[x_1]p[x_2]$, where $p$ is a path expression, $x_1$ and $x_2$ are terms, and $x_1 \not \in Str$.\\
A conjunction of path and built-in atoms $C = [X_1]p_1[Y_1] \cap \dots \cap [X_n]p_n[Y_n] \cap U_1\theta_1 V_1 \cap \dots \cap U_k \theta_k V_k$ is said to be \emph{safe} if all variables in $C$ are \emph{range restricted}, i.e. if
\begin{itemize}
 	\item for every $[X_i]p_i[Y_i]$, either $X_i$ is a constant (node identifier of a string), or there is some $[X_j]p_j[Y_j]$ in $C$ where $X_j$ is range restricted;
    \item for every built-in term $U_i\theta_i V_i$ occurring in $C$, if $\theta_i$ is equal to $"="$ then at least one of the two terms is range restricted; otherwise both $U_i$ and $V_i$ must be range restricted.
 \end{itemize}

A rootless tree formula is an expression of the form $p(\Phi_1 \land \dots \land \Phi_k)$ where $\Phi_i$ is a rootless path formula (expression of the form $p[y]$ where $p$ is a path expression and $y$ is a term) or a rootless tree formula and $p$ is a path expression. A tree atom is an expression of the form $[x]T$ where $T$ is a rootless tree formula and $x$ is a term.

@@

\begin{example}
Consider the XML tree $XT$ from Example \ref{example1ref}. Thereafter the integrity constraint that there must exist at least two books differents titles is expressed as

\begin{align*}
\forall(X)&[[root]/bib[X] \supset\\
&\exists(Z_1, Z_2)([X]/book/title/S[Z_1] \land [X]/book/title/S[Z_2] \land Z_1 \not = Z_2)]
\end{align*}\qed
\end{example}

@@

\begin{define}[Functional Dependency]\label{fd1}
Being given a DTD $D$, a {\sl functional dependency} on $D$ is an expression of the form $S \rightarrow p$, where $S$ is a finite non empty subset of $paths(D)$ and $p \in paths(D)$.\qed
\end{define}

@@

The main difference between two definitions of the functional dependency is that the former definition can use any path from $paths(D)$, whereas the latter considers that each $P_i (i \in [1,n])$ is from $StrPaths(D)$. That means the constraint from Example \ref{fdExample} cannot be expressed by functional dependency defined in \ref{fd2}, because of $/bib/book/written\_by/author$ path. Nevertheless, modified constraint from Example \ref{fdExample} expressed by FD defined in \ref{fd2} is shown in Example \ref{fd2Example}.

\begin{example}\label{fd2Example}
Consider the constraint $C$ saying that two distinct authors with different names of the same book cannot have the same value of attribute \texttt{ano}. The functional dependency expressing $C$ is defined as follows:

$$(bib/book, written\_by/author, (@ano \rightarrow name))$$\qed
\end{example}
